---
title: "STAT/MATH 495: Problem Set 07"
author: "Syed Abbas Shah"
date: "2017-10-24"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.width=8, fig.height=4.5, message=FALSE, warning = FALSE
  )
set.seed(76)

# Load packages
library(tidyverse)
library(broom)
library(knitr)
library(mosaic)
library(ROCR)

train <- read_csv("data/cs-training.csv") %>% 
  rename(Id = X1)
test <- read_csv("data/cs-test.csv") %>% 
  rename(Id = X1)
submission <- read_csv("data/sampleEntry.csv")
```

Information on the competition can be found [here](https://www.kaggle.com/c/GiveMeSomeCredit/data).



# Collaboration

Please indicate who you collaborated with on this assignment: 

No one.

#Exploratory analysis:

First, let's look at the response variable:
```{r}
prop.table(table(train$SeriousDlqin2yrs))
```
The vast majority of people in the dataset did not suffer from Serious Delinquency. Only about 6.7% of the people included did.

Now we can see how the response variable relates to some of the covariates of interest.
```{r EDA Age, message=FALSE, warning=FALSE}
train$age1 <- cut(train$age, breaks=c(25,30,35,40,45,50,55,60,65,70,75,80,85,90,100 ))
table1<-table(train$age1,train$SeriousDlqin2yrs)
x<-prop.table(table1,1)
kable(x, format="html", caption = "Percentages of Serious Delinquency in different Age groups")
plot(x)
```
There seems to be a pattern here. The risk of serious delinquency seems to be declining as age increases. 



```{r EDA monthly income, echo=FALSE, message=FALSE, warning=FALSE}
train$mincome1 <- cut(train$MonthlyIncome, breaks=c(0,2500,5000,7500,10000,150000,100000,3100000))
table1<-table(train$mincome1,train$SeriousDlqin2yrs)
x<-prop.table(table1,1)
kable(x, format="html", caption = "Percentages of Serious Delinquency in different Income groups")
plot(x)
```
Similarly, the incidents of Serious Delinquency decrease as Monthly Income increases.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
train$dratio1 <- cut(train$DebtRatio, breaks=c(0,.15,.3,.45,.6,.75,1,100,350000))
table1<-table(train$dratio1,train$SeriousDlqin2yrs)
x<-prop.table(table1,1)
kable(x, format="html", caption = "Percentages of Serious Delinquency in different Age groups")
plot(x)
```

The incidents of Serious Delinquents seem to increase for higher values of debt ratio, but decrease again in the right-most categories.


# Build binary classifier

I'm going to stick with Age as the predictor because the pattern was the most clearly discernable in my Exploratory Data Analysis.


```{r}
model_formula <- as.formula(SeriousDlqin2yrs~age)
model_logistic <- glm(model_formula, data=train, family="binomial")

model_logistic %>% 
  broom::tidy(conf.int=TRUE)
```


```{r, message=FALSE, warning=FALSE}
fitted_model <- model_logistic %>% 
  broom::augment() %>% 
  as_tibble() %>% 
  mutate(p_hat = 1/(1 + exp(-.fitted)))
predictions <- model_logistic %>% 
  broom::augment(newdata=test) %>% 
  mutate(p_hat = 1/(1 + exp(-.fitted)))
# Logistic regression is fitted in log-odds(p) space
ggplot(NULL) +
  geom_line(data=fitted_model, aes(x=age, y=.fitted), col="blue") +
  geom_point(data=predictions, aes(x=age, y=.fitted), col="red") +
  labs(x="Age", y="Fitted log-odds of p_hat", title="Fitted log-odds of probability of Having Serious Delinquency vs Age")
```



```{r, message=FALSE, warning=FALSE}
x<-ggplot(NULL) +
  geom_jitter(data=fitted_model, aes(x=age, y=SeriousDlqin2yrs), height=0.1, alpha=0.05) +
  geom_line(data=fitted_model, aes(x=age, y=p_hat), col="blue") +
  geom_point(data=predictions, aes(x=age, y=p_hat), col="red") +
  labs(x="Age", y="p_hat", title="Fitted probability of experiencing Delinquency in 90 days  vs Age")
x 
```




# ROC curves for Model + Random guessing:

Based on the ultimate classifier you choose, plot a corresponding ROC curve.

```{r, message=FALSE, warning=FALSE}
profiles_train_augmented <- model_logistic %>% 
  broom::augment() %>% 
  as_tibble() %>% 
  mutate(p_hat = 1/(1+exp(-.fitted)))
pred <- prediction(predictions = profiles_train_augmented$p_hat, labels = profiles_train_augmented$SeriousDlqin2yrs)
perf <- performance(pred, "tpr","fpr")
auc <- as.numeric(performance(pred,"auc")@y.values)
auc
plot(perf, main=paste("Area Under the Curve =", round(auc, 3)))
abline(c(0, 1), lty=2)
```


The dark line shows the area under the curve for our model, whereas the diagonal line represents the model which guesses randomly (50-50%). We can see that our model is slightly better than guessing randomly.



#Submission

```{r, message=FALSE, warning=FALSE}
submission$Probability <- predictions$p_hat
write_excel_csv( submission, "submission.csv")
```

